---
title: "Group Lab Exponential"
author: "Hamiz Mirza"
date: "2023-04-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
library(tsibble)
library(ggplot2)
library(ggfortify)
library(fable)
```

(a)  Create a vector of 1000 white noise values wt which are iid normal with mean 0 and sd 5.

```{r}
set.seed(100)
w <- rnorm(1000, mean = 0, sd = 5)
y <- rep(0,1000)
for (tt in 2:1000){
  y[tt] <- y[tt-1] + w[tt-1]
}

rw <- tsibble(tt = 1:1000, y, index = tt)
rw
autoplot(rw)
acf(rw)

rw %>% model(NAIVE(y)) %>% report()

```

2. Exponential smoothing state space model

(a)  With the same white noise values wt you created in question 1, and smoothing parameter α = 0.5, use the state equation lt = lt−1 + α*wt−1 to create a vector of levels lt.
(b)  Use the observation equation yt+1 = lt + wt to create a time series.
(c)  Make your time series yt into a tsibble and plot it as well as its autocorrelogram.



```{r}
set.seed(101)
w <- rnorm(1000, mean = 0, sd = 5)
l_t <- rep(0,1000)
for (tt in 2:1000){
  l_t[tt] <- l_t[tt-1] + 0.5*w[tt-1]
}

ew <- tsibble(tt = 1:1000, l_t, index = tt)
autoplot(ew)

y_t <- rep(0,1000)
for(tt in 1:999){
  y_t[tt+1] <- l_t[tt] + w[tt]
}
rw2 <- tsibble(tt = 1:1000, y_t, index = tt)
autoplot(rw2)

```


(d)  Fit an ETS model, and check that the fitted values of the smoothing parameter α and variance σ2 are close to the correct values.

```{r}
fit <- rw2 %>% model(ETS(y_t ~ error("A")))
report(fit)
```

3. Repeat question 2 with α = 0.01. Compare the plots for all three questions.


4. Holt’s exponential smoothing with linear trend

(a)  With the same white noise values wt you created in question 1, and smoothing parameters

α = 0.5 and β = 0.5, use the state equations

yt =lt−1 +bt−1 +wt

lt =lt−1 +bt−1 +αwt

bt = bt−1 + αβwt

to create a simulated time series yt.

(b)  Make your time series yt into a tsibble and plot it.

(c)  Fit an ETS model, and check that the fitted values of the smoothing parameters αˆ, βˆ,and variance σˆ2 are close to the correct values.

5. Seasonal, multiplicative, and damping ETS models

(a)  Using the AirPassengers data, fit a simple exponential smoothing model ETS(A,N,N), make a 4-year forecast, and plot the forecast and data on the same chart.

(b)  Repeat this for other ETS models, including Holt-Winters ETS(A,A,A), multiplicative Holt-Winters ETS(M,A,M), and investigate the effect of damping on the trend term.





