---
title: 'Group Lab 5: Regression'
author: "Hamiz Mirza"
date: "2023-03-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fpp3)
library(tsibble)
library(ggfortify)
library(dplyr)
library(tidyr)
```

1. Get started by converting to daily data. Use index by(Date) and then summarize to get the total Demand for each day and the daily high temperature.

```{r}
dailydata <- vic_elec %>%
  index_by(Date) %>%
  summarize(Total_Demand = sum(Demand), Daily_High = max(Temperature))
```

2. Make a seasonal plot of Demand (with gg season) to see the annual pattern of electricity use.

```{r}
gg_season(dailydata, Total_Demand)
```
3. Since electricity use depends on heating, cooling, and the length of days, we concentrate only on the months June, July, August (winter in Australia).
Filter the data to just select those three months and just the year 2013. Make time plots of Demand and daily high temperature.

```{r}
winterdata <- dailydata %>% filter_index("2013-06-01" ~ "2013-08-31")
ggplot(winterdata) + 
  aes(Date,Total_Demand) +
  geom_line()
```

```{r}
ggplot(winterdata) + 
  aes(Date,Daily_High) +
  geom_line()
```
4. Make a plot showing Demand as a function of daily high temperature. What do you observe?

```{r}
ggplot(winterdata, aes(x = Daily_High, y = Total_Demand)) +
  geom_point()
```
It is difficult to definitively discuss a relationship without conducting further analysis, however just from looking at the graph it appears that there is some confounding variable causing two seperate sets of data, both of which have a negative relationship. 

5. There are clearly two types of day in this data - days with and without work. Create a new variable that distinguishes the two types of day. Recreate your plot from (4) and color the points according to your new variable. Add regression lines for each group.

```{r}
winterworkdata <- winterdata %>% mutate(Weekend = if_else(weekdays(Date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
ggplot(winterworkdata, aes(Daily_High,Total_Demand, color = Weekend)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
6. Create a time series linear model of Demand using high temperature and your workday variable. Check the coefficients. Describe how demand changes with a 1◦C change in temperature. How much does a workday affect Demand?

```{r}
demandmodel <- winterworkdata %>% 
  model(TSLM(Total_Demand ~ Daily_High + Weekend))
report(demandmodel)
```
A decrease of 1 degrees Celsius results in an decrease of 3288.3  MW energy used. The weekend consumes significantly less energy, 36515.8 MW less energy compared to weekdays. 

7. Make a plot showing Demand as well as your fitted model. (You’ll need to augment() your model.)

```{r}
winterworkdataaugment <- augment(demandmodel, newdata = winterworkdata)
ggplot(winterworkdataaugment, aes(x = Date, y = Total_Demand)) +
  geom_line() +
  geom_line(aes(y = .fitted), color = 'red')
```
8. Plot the residuals of your model. What do you observe? Check if they are white noise using a Ljung-Box test.

```{r}
gg_tsresiduals(demandmodel)
```

The series sst.csv contains weekly sea surface temperatures in four regions of the Southern Pacific ocean, and is an indicator of the weather phenomenon known as El Nino.

```{r}
seatemp <- read.csv('https://mathstat.slu.edu/~clair/timeseries/data/sst.csv') 
sea <- seatemp %>%
  mutate(Date = yearweek(week))
```

1. Make a time plot of the Nino4 region temperatures and show the linear regression line with time as the predictor variable.

```{r}
sea4 <- sea %>% 
  as_tsibble(index = Date) %>%
  select(Date,Nino4)

ggplot(sea4) +
  aes(x = Date, y = Nino4) +
  geom_line() + 
  geom_smooth(method = "lm")
```
2. Fit a linear model for the Nino4 region temperatures with time as the predictor. What does the trend coefficient tell you about ocean temperatures in this part of the ocean?

```{r}
seamodel <- sea4 %>% model(TSLM(Nino4 ~ trend()))
report(seamodel)
```
3. Fit a linear model and include seasonal dummy variables. How many dummy variables did this require? Plot the fitted values and the data on the same chart.

```{r}
seamodel2 <- sea4 %>% model(TSLM(Nino4 ~ trend() + season()))
report(seamodel2)

augment(seamodel2) %>%
  ggplot(aes(x = Date)) +
  geom_line(aes(y = Nino4, color = 'blue')) +
  geom_line(aes(y = .fitted, color = 'red'))

glance(seamodel2)
```
51 dummy variables are required as we need the number of weeks - 1.

4. Fit a linear model and include Fourier series terms. Minimize cross validation error (CV) to select the best number of terms. Plot the fitted values and the data on the same chart.

```{r}
sea_fourier <- sea4 %>% model(TSLM(Nino4 ~ trend() + fourier(K = 2)))
report(sea_fourier)

augment(sea_fourier) %>%
  ggplot(aes(x = Date)) +
  geom_line(aes(y = Nino4, color = 'blue')) +
  geom_line(aes(y = .fitted, color = 'red')) 

glance(sea_fourier)
```
K = 2 minimizes the Cross-Validation error. We tested this by utilizing multiple values of K and finding out which one produced the lowest CV error.

5. How does your Fourier series model compare to the one with seasonal dummy variables?

The Fourier series model is more accurate when examining the Cross-Validation values compared to the Cross-Validation values of the model which utilized dummy variables.